Formal Synthesis of Digital Controllers for Observable Continuous Plants with Transient Performance Specifications

### 2018-01-31 TACAS2018/paper.tex
forall

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  However, in
order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and
$\mat{L}=\mat{L}_{\langle C \rangle}$, we must be able to ensure that
certain properties of the plant hold at any given real time.  In order to do
so, we accelerate the dynamics of the plant according to the model semantics
in Sec.~\ref{sec:model_semantics} and compute the state $\vec{x}_T$ at a
given time $T$,
%
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time
$t=T$.

@@

Next, we are interested in studying models with parametric inputs, which are
described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where
$\forall t\geq 0, \vec{u}(t)=\vec{u}$.  For such models, a state $\vec{x}_T$
is computed next.
%
\begin{lemma}
%
The state of the model
$\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a
parametric input $\vec{u}$ is
%
\begin{align}
\vec{x}_T&=\vec{x}(t=T)=\mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u} \text{ where } \mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B} 
\label{eq:continuous_tube_param}
\end{align}
%
where $\forall t \leq T,\ \vec{u}(t)=\vec{u}$, and $\mat{A}_T$ as in
\eqref{eq:continuous_tube_dyn2} \footnote{When $\mat{A}$ is not invertible,
we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\
0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$}. 
If $\mat{J}$ is diagonal and given a sampling time $T_d$ such that
$A_d=e^{\mat{A} T_d}$, then
$\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}_d\vec{u})
\text{ with } T \in \mathbb{R}_0^+$ (where $\mat{A}_d^{\frac{T}{T_d}} =
e^{log(\mat{A}_d) \frac{T}{T_d}} = e^{\mat{A} T}$) is the state of the
system at time $T$.
%
\end{lemma}

@@

Gershgorin disks~\cite{van1996matrix} allow us to limit the error in the
expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in [1, ...\,,n],
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are
the true eigenvalues of the observer $\mat{A}_o$ and $\hat{\lambda}_{oi}$
the corresponding eigenvalues of the closed loop.  We note that in the case
of stability we only care that their norms are smaller than one, which is
equivalent to upper bounding the estimation of the largest eigenvalue. 
Hence, given a desired upper value $\overline{\lambda}$ (see Section
\ref{sec:transientspecification}), we apply Jury's criterion on the
polynomial
%
\begin{equation}
P_L'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of $\mat{A}_c$, the procedure is the same except that we need to look at the columns shared with $\mat{A}_c$.
Thus we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_K'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

A safety specification leads to a requirement on the states of the model. 
Consequently, the controller must ensure that the state never violates the
requirement.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}^n: = & \forall k \in [0, ... , n], \mat{G}\vec{x}_k \leq \vec{h} \text{ and } \underline{y} \leq y_k \leq \overline{y}\nonumber\\
\phi_\mathit{safety}: = & \phi_\mathit{safety}^\infty
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \eqref{eq:observer_LTI_FWL_cf}.

@@

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p],
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  For the proof see
Appendix~\ref{sec:proofs}
%
\end{theorem}

@@

 Let $\mat{J}$ be a diagonal matrix, such that 
 $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
 Let us now take a different sample rate $T_2$, such that 
 $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
 We can then say that 
 \begin{equation*}
 {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
 \end{equation*}
 This proves the theorem for $gm(\lambda_i)=1$.
 We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
 referenced to the original eigenvalues modified by the sampling rate.
  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix, then
\begin{align*}
 \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k\nonumber\\
 &=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
 \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
& &  &\lambda_i^k \\
\end{array} \right ] T_1^k
\end{align*}
The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
\begin{align*}
\forall j>i, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
\end{align*}
which completes the proof for $gm(\lambda_i)>1$. \qed
\end{proof}

@@

%-------------------------------
\subsubsection{Transient response} 
\label{sec:transientspecificationproof}
%-------------------------------
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $p^{th}$ order system $\lambda_i =e^{\omega_i}, i \in [1, ...\,,p]$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in [1, ...\,,p]
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.
\end{theorem}
\begin{proof}
In a SISO system, there are two values that typically define the
response-time of the system (using a step response).
%
\begin{enumerate}
\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step change in the input.
\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
\end{enumerate} 
%
%In the case of first and second order systems, the decay is governed by the equations 
%%
%\begin{align*}
%&y_{decay}=
%&\left\{
%\begin{array}{cc}
%e^{-|\omega| t}& 1^{st} order\\
%e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
%e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
%\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
%\end{align*}
%where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
%These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
%All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.

>>>>

The time discretization in Sec.~\ref{sec:model} only allows us to evaluate
the state of the plant at integer multiples of the sample time.  However, in
order to synthesize $\mat{K}=\mat{K}_{\langle C \rangle}$ and
$\mat{L}=\mat{L}_{\langle C \rangle}$, we must be able to ensure that
certain properties of the plant hold at any given point in time.  In order
to do so, we accelerate the dynamics of the plant according to the model
semantics in Sec.~\ref{sec:model_semantics} and compute the state
$\vec{x}_T$ at a given time $T$,
%
% namely we find an equation for $\tau^n(\vec{x},\vec{u})$ given $\tau(\vec{x},\vec{u})$. 
%
namely the state of the system $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)$ at time
$t=T$.

@@

Next, we are interested in studying models with parametric inputs, which are
described by $\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$, where
$\forall t\geq 0.\,\vec{u}(t)=\vec{u}$.  For such models, a state $\vec{x}_T$
is computed next.
%
\begin{lemma}
%
The state of the model
$\dot{\vec{x}}(t)=\mat{A}\vec{x}(t)+\mat{B}\vec{u}(t)$ at time $T$ given a
parametric input $\vec{u}$ is
%
\begin{align}
\vec{x}_T&=\vec{x}(t=T)=\mat{A}_T\vec{x}_0 + \mat{B}_T\vec{u} \text{ where } \mat{B}_T=\mat{A}^{-1}(\mat{I}-\mat{A}_T)\mat{B} 
\label{eq:continuous_tube_param}
\end{align}
%
where $\forall t \leq T.\, \vec{u}(t)=\vec{u}$, and $\mat{A}_T$ as in
\eqref{eq:continuous_tube_dyn2} \footnote{When $\mat{A}$ is not invertible,
we use $\left[\begin{array}{cc}\mat{A}_T&\mat{B}_T\\
0&\mat{I}\end{array}\right]=e^{\left[\begin{array}{cc}\mat{A}&\mat{B}\\0&0\end{array}\right]}$}. 
If $\mat{J}$ is diagonal and given a sampling time $T_d$ such that
$A_d=e^{\mat{A} T_d}$, then
$\vec{x}_T=A_d^{\frac{T}{T_d}}\vec{x}_0+(\mat{I}-\mat{A}_d^{\frac{T}{T_d}}\mat{B}_d\vec{u})
\text{ with } T \in \mathbb{R}_0^+$ (where $\mat{A}_d^{\frac{T}{T_d}} =
e^{log(\mat{A}_d) \frac{T}{T_d}} = e^{\mat{A} T}$) is the state of the
system at time $T$.
%
\end{lemma}

@@

Gershgorin disks~\cite{van1996matrix} allow us to limit the error in the
expected eigenvalues.  Let $\delta_o=\max(\norm{{\mat{A}_e}_{i,*}}_1)$
(\emph{i.e.} the largest row norm in $\mat{A}_e$) be the largest error bound
for any eigenvalue in $\mat{A}_e$.  Then $\forall i \in \{1, \ldots,n\},
|{\lambda_o}_{i}-\hat{\lambda}_{oi}|<\delta_o$, where ${\lambda_o}_{i}$ are
the true eigenvalues of the observer $\mat{A}_o$ and $\hat{\lambda}_{oi}$
the corresponding eigenvalues of the closed loop.  We note that in the case
of stability we only care that their norms are smaller than one, which is
equivalent to upper bounding the estimation of the largest eigenvalue. 
Hence, given a desired upper value $\overline{\lambda}$ (see Section
\ref{sec:transientspecification}), we apply Jury's criterion on the
polynomial
%
\begin{equation}
P_L'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_o)\overline{\lambda} \text{ and } c_j=l_j-a_j.
\end{equation}
%
In the case of $\mat{A}_c$, the procedure is the same except that we need to look at the columns shared with $\mat{A}_c$.
Thus we obtain our bound $\delta_c=\max(\norm{{\mat{A}_e}_{*,i}+{\mat{B}_e}_{*,i}}_1)$
(\emph{i.e.} the largest column norm in $ \left[\begin{array}{c}\mat{A}_e\\ \mat{B}_e\end{array}\right]$) and
apply Jury's Criterion over
%
\begin{equation}
P_K'(z)=\sum_{j=0}^p c_j\overline{\lambda'}^{p-j}z^j, \text{ where }  \overline{\lambda'}=(1+\delta_c)\overline{\lambda} \text{ and } c_j=k_j-a_j.
\end{equation}

@@

A safety specification leads to a requirement on the states of the model. 
Consequently, the controller must ensure that the state never violates the
requirement.  In~this work, the safety property is expressed as:
%
\begin{align}
\phi_\mathit{safety}^n: = & \forall k \in \{0, \ldots , n\},
  \mat{G}\vec{x}_k \leq \vec{h} \text{ and } \underline{y} \leq y_k \leq \overline{y}\nonumber\\
\phi_\mathit{safety}: = & \phi_\mathit{safety}^\infty
\label{eq:safetyspec}
\end{align}
%
where $\mat{G}$ and $\vec{h}$ define a set of inequalities that describe 
a safe set, within the state-space $\mathbb{R}^p$.
%
Note that FWL effects are not explicitly captured here
as they are implicit from Equation \eqref{eq:observer_LTI_FWL_cf}.

@@

\begin{theorem}
%
Given a set of eigenvalues for a discrete time model of a $p^{th}$ order
stable system $\{\lambda_i =e^{\omega_i} \mid i \in [1, ...\,,p],
|\lambda_i| \leq 1 \}$, the settling time of the system is upper-bounded by
%
\begin{equation}
%
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }
  \overline{\lambda} = \max(|\lambda_i|) \text{ and }
  \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
%
\label{eq:set_time}
\end{equation}
%
where $0\leq P_s \leq 1$ is the desired portion of the target within which
the output must remain after $t_s$.  For the proof see
Appendix~\ref{sec:proofs}
%
\end{theorem}

@@

 Let $\mat{J}$ be a diagonal matrix, such that 
 $${\lambda_1}_i=e^{\lambda_i T_1}, \forall \lambda_i \in \mat{J}.$$
 Let us now take a different sample rate $T_2$, such that 
 $${\lambda_2}_i=e^{\lambda_i T_2}, \forall \lambda_i \in \mat{J}.$$
 We can then say that 
 \begin{equation*}
 {\lambda_2}_i=e^{\lambda_i T_1 \frac{T_2}{T_1}}={\lambda_1}_i^{\frac{T_2}{T_1}} \Rightarrow A_2=A_1^{\frac{T_2}{T_1}}.
 \end{equation*}
 This proves the theorem for $gm(\lambda_i)=1$.
 We now note that given an exponentiated Jordan form with geometric multiplicity, the upper diagonal terms can be
 referenced to the original eigenvalues modified by the sampling rate.
  Let $\mat{J}\in \mathbb{R}^{p \times p}$ be a Jordan Canonical matrix, then
\begin{align*}
 \mat{J}_1&=\sum_{k=0}^\infty \frac{1}{k!}\left(\mat{J}T_1\right)^k\nonumber\\
 &=\sum_{k=0}^\infty \frac{1}{k!} \left [ \begin{array}{cccc}
 \lambda_i^k  & \binom{k}{1}  \lambda^{k-1} & \hdots  & \binom{k}{p-1} \lambda_i^{k-p+1} \\
& \lambda_i^k  & \binom{k}{1}  \lambda_i^{k-1} & \vdots \\
\vdots & \ddots & \ddots & \vdots \\
& &  &\lambda_i^k \\
\end{array} \right ] T_1^k
\end{align*}
The eigenvalues remain the same as in the diagonal case, but the upper triangular terms are of the form:
\begin{align*}
\forall j>i.\, c_{ij}&=\sum_{k=j-i}^\infty \frac{1}{k!}\binom{k}{j-i} \lambda_i^{k-(j-i)}T_1^k\\
&=\frac{1}{(j-i)!}\sum_{k=j-i}^\infty \frac{1}{(k-(j-i))!} \lambda_i^{k-(j-i)}T_1^k\nonumber\\
&=\frac{T_1^{j-i}e^{\lambda_i T_1}}{(j-i)!}\nonumber
\end{align*}
which completes the proof for $gm(\lambda_i)>1$. \qed
\end{proof}

@@

%-------------------------------
\subsubsection{Transient response} 
\label{sec:transientspecificationproof}
%-------------------------------
\begin{theorem}
Given a set of eigenvalues for a discrete time model of an $p^{th}$ order system $\lambda_i =e^{\omega_i}, i \in [1, ...\,,p]$, the settling time of the system is upper-bounded by
\begin{equation}
t_s<-log_{\overline{\lambda}}({P_s}), \text{ where }  \overline{\lambda} = \max(|\lambda_i|) \text{ and } \overline{\lambda}<1,\ \forall i \in \{1, \ldots,p\}
\label{eq:set_time}
\end{equation}
where $0\leq P_s \leq 1$ is the desired percentage of the target within which the output must remain after $t_s$.
\end{theorem}
\begin{proof}
In a SISO system, there are two values that typically define the
response-time of the system (using a step response).
%
\begin{enumerate}
\item The \emph{rise time} is defined as the time it takes the output to reach a percentage (eg $90\%$) of the target value given a step change in the input.
\item The \emph{settling time} is defined as the time it takes the output to stabilize within a margin of the target value (eg $10\%$).
\end{enumerate} 
%
%In the case of first and second order systems, the decay is governed by the equations 
%%
%\begin{align*}
%&y_{decay}=
%&\left\{
%\begin{array}{cc}
%e^{-|\omega| t}& 1^{st} order\\
%e^{\zeta |\omega| t}\left(\frac{\sin\left(|\omega| t\sqrt{1-\zeta^2}+\sin^{-1}\sqrt{1-\zeta^2}\right)}{\sqrt{1-\zeta^2}} \right)&\zeta<1\\
%e^{-|\omega| t}-|\omega| t e^{-|\omega| t}&\zeta=1\\
%\frac{|\omega| e^{-|\omega_2| t}-|\omega_2| e^{-|\omega| t}}{|\omega|-|\omega_2|}&\zeta>1\end{array}\right.
%\end{align*}
%where $|\omega|$ and $|\omega_2|$ are the magnitudes of the eigenvalues (roots) of the system and  $\zeta=-\cos(\theta)$ where $\theta$ is the polar angle of the conjugate pair if it exists ($\zeta \geq 1$ if it doesn't).
%These equations become increasingly complicated for higher order systems. However, we note their strong dependency on the magnitude of the eigenvalues, and use this property to define an upperbound on the decay function.
%All of the above equations will decay faster than $e^{-|\omega| t}$ given $|\omega| < |\omega_2|$, hence we may use this function as our upper bound.
